Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-04 paper_jasa.tex
updated comments

The experiment is designed to study the ability of human observers to detect the effect of a single variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical UMP test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. %\green{Explain why this was used.} \red{There was not any real reason for this. It was just arbitrarily chosen to generate some data. When I first started to think of applying visual inference in regression setting, I was thinking of a dataset (I found in a book) of weight of babies by their gestational age (in week) where author wanted to test the difference of mean weight for gender.   Thus for our experiment I generated values similar to a variable like age. the discrete variable was thought similar to gender.}
Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual ydata plot was generated by making a side-by-side boxplot of the residuals (Table\ref{tbl:stat_multiple} case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2$, and these were plotted in the same way as side-by-side boxplots. The actual data plot was randomly placed among these null data plots to produce the lineup. Figure \ref{fig:test_category} is an example of one of these lineups. The actual data plot is number 13 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.  \green{I think it is plot11, not 13!!!!!!!! Please check!}

@@

The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power. Suppose $\gamma$ denotes the classical UMP test power for each parameter combination shown in table \ref{tbl:experiment_params}. Since the expected power of visual inference is very close to the power of classical UMP test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$. For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.1. Thus we have $ME =1.96 \sqrt{ \frac 1 n_{\gamma} \gamma(1-\gamma)} \le 0.05$ \green{0.05 or 0.1???} which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$

>>>>

The experiment is designed to study the ability of human observers to detect the effect of a single variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical UMP test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. %\green{Explain why this was used.} \red{There was not any real reason for this. It was just arbitrarily chosen to generate some data. When I first started to think of applying visual inference in regression setting, I was thinking of a dataset (I found in a book) of weight of babies by their gestational age (in week) where author wanted to test the difference of mean weight for gender.   Thus for our experiment I generated values similar to a variable like age. the discrete variable was thought similar to gender.}
Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual ydata plot was generated by making a side-by-side boxplot of the residuals (Table\ref{tbl:stat_multiple} case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2$, and these were plotted in the same way as side-by-side boxplots. The actual data plot was randomly placed among these null data plots to produce the lineup. Figure \ref{fig:test_category} is an example of one of these lineups. The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.

%\green{I think it is plot11, not 13!!!!!!!! Please check!}

@@

The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power. Suppose $\gamma$ denotes the classical UMP test power for each parameter combination shown in table \ref{tbl:experiment_params}. Since the expected power of visual inference is very close to the power of classical UMP test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$. For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05. Thus we have $ME =1.96 \sqrt{ \frac 1 n_{\gamma} \gamma(1-\gamma)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
